---
title: "Base R"
#author: "Dookyung Kim"
#date: "2018-11-13"
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
mainfont: NanumGothic
---
exclude: true

```{r, message=FALSE, warning=FALSE, include=FALSE}
options(
  htmltools.dir.version = FALSE, # for blogdown
  width=80
#  repos=structure(c(CRAN="http://cran.r-project.org"))
)

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# library(emo)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```
---
## R as a Calculator

.pull-left[
#### Basic calculation
```{R }
2+2
2^2
(1-2)*3
1-2*3
```

#### Built-in Functions
```{R }
sqrt(2)
sin(pi)
exp(1)
log(10)
log(10, base=10)
```
]

.pull-right[
#### Variable Assignment
```{r error=TRUE}
x = 2     # assignment
x+3

pi        # pi is a built-in constant

e^2       # Error

e = exp(1)
e^2

(x <- 'apple')   # print
```

#### Getting Help
```{r eval=FALSE}
# 특정함수에 대한 도움말 보기
?mean
# 특정 용어"에 대한 도움말 검색
help.search('weighted mean')
# 특정 패키지"에 대한 도움말 보기
help(package = 'deplyr')        
```
]

---
## (Almost) Everything is a Vector

.pull-left[
#### Creating Vectors
```{r}
# Join elements
c(2, 4, 6)

# An interger sequence
2:6              
(2:6) * 0.3

# A complex sequence
seq(2, 3, by=0.5)

# Repeat a vector
rep(1:2, times=3)

# Repeat elements of a vector
rep(1:2, each=3)  

rep(1:4, c(2,1,2,1))
rep(1:4, each=2, len=4)
```
]

.pull-right[
#### Vector function
```{r}
x <- c(1, 3, 5, 2, 1, 5, 4, 6, 2, 5)

# Return x sorted
sort(x)

# Return x reversed
rev(x)

# See counts of values
table(x)

# See Unique values
unique(x)
```
]
---
## Maths Functions works for vectors

```{r}
x <- c(1,2,3,4,5)
y <- c(4,5,3,5,9)

log(x)
sum(x)
exp(x)
round(exp(x), 2)
mean(x)
median(x)
max(x)
min(x)
var(x)
cor(x, y)
sd(x)
rank(x)
```


---
## Types of vectors

The fundamental building block of data in R are vectors. R has two fundamental vector classes:

- **atomic vectors** : *same* type elements 
```{r}
c(3, 4, 5)
c("apple", "banana", "monkey")
```

- **Lists** : collections of *any* type of R object.

```{r}
l <- list(3, list("apple","banana"), FALSE, 3.35)
l

```
---
## R has six atomic vector types 

*  logical  :  boolean values (TRUE or FALSE)      
*  double   :  2.44
*  integer  :  3L
*  character : "apple"
*  complex
*  raw 


### How to convert types

```{r}
a <- c(TRUE, FALSE, TRUE)
print(a)
typeof(a)

a <- as.numeric(a)
print(a)
typeof(a)

a <- as.logical(a)
print(a)
typeof(a)

a <- as.character(a)
print(a)
typeof(a)

a <- as.factor(a)
print(a)
typeof(a)
```

---
## Selecting Vector Elements

- By Position
```{r}
x <- c()
x <- 1:6

x[4]
x[-4]
x[2:4]
x[-(2:4)]
x[c(1,3,5)]
```


- By Value
```{r}
x[x==3]
x[x<3]
x[x %in% c(1,2,5)]
```
---
## Condtions

- a == b
- a != b
- a > b
- a < b
- a >= b
- a <= b
- is.na(a)
- is.null(a)

---

## Comparisons

  Operator  |  Comparison                |  Vectorized?
:-----------|:---------------------------|:-----------------
 `x < y`    |  less than                 |  Yes
 `x > y`    |  greater than              |  Yes
 `x <= y`   |  less than or equal to     |  Yes
 `x >= y`   |  greater than or equal to  |  Yes
 `x != y`   |  not equal to              |  Yes
 `x == y`   |  equal to                  |  Yes
 `x %in% y` |  contains                  |  Yes (for `x`)




---
## Logical (boolean) operations

|  Operator  |  Operation    |  Vectorized? 
|:-----------|:--------------|:-------------
| <code>x &#124; y</code>    |  or           |   Yes        
| `x & y`    |  and          |   Yes        
| `!x`       |  not          |   Yes        
| <code>x &#124;&#124; y</code> |  or           |   No         
| `x && y`   |  and          |   No         
|`xor(x,y)`  |  exclusive or |   Yes        

---
---
## Vectorized?

```{r}
x = c(TRUE,FALSE,TRUE)
y = c(FALSE,TRUE,TRUE)
```

.pull-left[
```{r}
x | y
x || y
```
]

.pull-right[
```{r}
x & y
x && y
```
]


---
## handling data as a vector

```{r}
typos.draft1 = c(2,2,3,4,5,3)
typos.draft2 = typos.draft1  # make a copy
typos.draft2[1] = 0          # assign the first value to 0
typos.draft2
typos.draft2[2]
typos.draft2[4]

typos.draft2
typos.draft2[-4]
typos.draft2[c(1,2,5)]
typos.draft2 == 3

```
---
## Selecting some values with logical operators

```{r}
typos = c(2,3,0,3,1,0,0,1)
typos[typos==1]
typos[typos>1]
typos[typos<=1]
typos[typos>0 & typos<=2]
typos[typos<=0 | typos>2]
```
---
##  Matrix in R





```{r}
# Generating Matrices
matrix(1:8, ncol=4, nrow=2)
matrix(c(1,2,3,11,12,13), nrow=2, ncol=3, byrow=T)
matrix(c(1,2,3,11,12,13), nrow=2, ncol=3, byrow=TRUE,
       dimnames=list(c("row1","row2"),c("C.1","C.2","C.3")))

# Special Matrices
diag(3)
diag(4:6)
A = matrix(0,3,3)
A[upper.tri(A, diag=TRUE)] = 1:6
A 

# Generating sub-matrices and sub-vectors
A = matrix(1:9,3,3)
A
A[,1]
A[2,2]
A[1,c(2,3)]
A[1:2,2:3]
```

---
## Operations

```{r}
# Fundamental oprations of vectors
c(1,3,5) + c(3,5,7)
c(1,3,5) * c(3,5,7)
c(1,3,5) / c(3,5,7)
c(1,3,5) + 1
c(1,3,5) * 2

# Fundamental operations of matrices
A = matrix(1:9,3,3)
B = matrix(2:10,3,3)
A;B


A + B
A * B
A %*% B   # matrix multiplication
t(A)
solve(A*B)
solve(A*B) %*% (A*B)

# Transformation of vectors and metrices
A = matrix(1:4,2,2);A
is.vector(A); is.matrix(A)
as.vector(A)
x = 1:3; y= 11:13
cbind(x,y)
rbind(x,y)
```
---
## Data Frames

```{r error=TRUE}
# Data frames and matrices
A = matrix(1:4,2,2)
dA = data.frame(A)
dA
is.matrix(dA)
is.data.frame(dA)
dA * dA

dA %*% dA    # does not work
A %*% A

# imported data set is a data frame
dataset = read.csv("./data/data.csv", header=T, stringsAsFactors=F)
head(dataset,5)
is.data.frame(dataset)
is.matrix(dataset)
dataset[,1]
dataset$AGE
dataset$SEX

```

---
## more about an object

```{r}
data(iris)
iris
str(iris)
class(iris)
head(iris)
```



---
## List

```{r}
# Collections of vectors and matrices
x1 = c(1,2,3)
x2 = c(4,5,6,7) 
x3 = c(7,8,9,10,11)
list(x1,x2,x3)
list(x1,x2,x3)[[1]]
list(a=x1,b=x2,c=x3)
list(a=x1,b=x2,c=x3)$a
```


```{r}
l <- list(x = 1:5, y = c('a', 'b'))
          
l[[2]]
l[1]
l$x
l['y']

```
---

##  controls

```{r}
# if statements 
x = -1
if(x>0){y="positive"}
y

if(x>0){y="positive"}else{y="nagative"}
y

ifelse(x>0, "positive","negative")
```

---

```{r}
# for loops
a=b=c()
for (i in 1:5){
    a[i] = i
    b[i] = a[i] + i^2
}

a
b

M = matrix(NA,3,5)
for (i in 1:3){
    for(j in 1:5){
        M[i,j] = i + j^3
    }
}

M
```

---

```{r}
# While and Repeat loops
i=0; a=100
while(a[i+1]>0.01){
    i=i+1
    a[i+1] = a[i]/10
}
i
a

i=0;a=100
repeat{
    i=i+1
    a[i+1]=a[i]/10
    if(a[i+1]<=0.01) break
}
i
a

```
---

## Graph in R

```{r}
#Histogram
x = rnorm(10000)
hist(x, prob=T, main="H", xlab="X", ylab="Y")
hist(x, prob=T, breaks=10, xlim=c(-4,4), ylim=c(0,0.4))
hist(x, prob=T, breaks=20, xlim=c(-4,4), ylim=c(0,0.4))
hist(x, prob=T, breaks=30, xlim=c(-4,4), ylim=c(0,0.4))
v=seq(-5,5,0.01)
lines(v,dnorm(v))
lines(v,dnorm(v), col="blue")
lines(v,dnorm(v), col="blue",lwd=4)
```
---
```{r}
# Boxplots
x = runif(10000, min=0, max=1)
boxplot(x, main="Boxplot")

head(OrchardSprays)
boxplot(decrease ~ treatment, data=OrchardSprays, col="bisque")
```


---
```{r}
# plot functions
x = seq(-3,3,0.25)
y = dnorm(x)
plot(x,y)

plot(x,y,type="l")

plot(x,y,type="h")
points(x,y,pch=2)
lines(x,y,lty=2)

```

---
##  Random number generations

```{r}
x = rnorm(100000, mean=5, sd=5)
hist(x, main="N(5,25)")
abline(v=5, lty=2,lwd=3)

y=rgamma(10000,shape=5,scale=5)
hist(y,main="Gamma(5,5)")
```
---

##  functions in R

```{r}
f1 = function(x,y,z) {
    x^2+log(y)+exp(z)
}
f1(2,3,4)

f2 = function(A,B){
    temp = A %*% B
    det(temp)
}

A0 = matrix(1:4,2,2)
B0 = matrix(5:8,2,2)
f2(A0,B0)
```


---

## Install Packages

```{r}
# installing R packages
# install.packages("ggplot2")

# importing R packages
library(ggplot2)
```
---
## Working Directory

```{r}
getwd()
#setwd('C://file/path')

```

---
## Environment

```{r}
rm(list = ls())
a <- "apple"
b <- c(1,2,3)
c <- 1

ls()
rm(a)
ls()
rm(list = ls())
```


---
## Reading a text file

```{r}
# read csv
read.csv("./data/data.csv", header=T, stringsAsFactors=F)

# text file
read.table("./data/data.txt")
read.table("./data/data.txt", header=T, sep="|", stringsAsFactors=F)
```

---

## Read & Write data

```{R eval=FALSE}
## create txt file
# fileConn<-file("file.txt")
# writeLines(c("Hello","World"), fileConn)
# close(fileConn)

df <- c("Hello","World")
write.table(df, 'file2.txt')

df2 <- read.table('file2.txt')
print(df2)
```

```{R eval=FALSE}
df <- c("apple","graph")
write.csv(df, 'file3.csv')

df3 <- read.csv('file3.csv')
print(df3)
```

```{R eval=FALSE}
df <- c("apple3","graph3")
save(df, file = 'file4.Rdata')

load('file4.Rdata')
print(df)
```


---
## Reference:

- Ruey S. Tsay, An Introduction to Analysis of Financial Data with R
- Taeyoung Park, Bayesian Inference & Computation